On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-20 paper.tex
Used d1

We need one additional technical definition. We call an ordered triple of roots $(\alpha, \beta, \gamma) \in \Phi \times \Phi \times \Phi$ an {\it $\rA_3$-triple}
 if these vectors form a basis of a root subsystem $\Psi \subseteq \Phi$ of type $\rA_3$, and moreover if the Dynkin diagram of $\Psi$ with respect to this basis has the form
 $\xymatrix{\circ_\alpha \ar@{-}[r] & \circ_\beta \ar@{-}[r] & \circ_\gamma}.$ %TODO: Draw better diagram :)

@@

For $m\geq 2$ define $G_m$ as the group presented by generators $\mathcal{X}_m$ and the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{\alpha}(\xi) x_{\alpha}(\eta) & = x_{\alpha}(\xi+\eta),&  \xi,\eta\in B_d,\ d\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\beta}(\eta))] &  = 1, & \text{in the case $\alpha+\beta\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_d,\ \eta \in B_e,\ d,e\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\beta}(\eta)] & = x_{\alpha+\beta}(N_{\alpha,\beta}\xi\eta), & \text{in the case $\alpha+\beta\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_d,\ \eta \in B_e,\ d, e, d+e \leq m.
 \end{align}
For $m=1$ define $G_1$ as the group presented by generators $\mathcal{X}_m$, three families of relations $(a_1), (b_1), (c_1)$ as above and the following additional family of relations:
\begin{align}  \label{eq:d1} \tag{$d_1$} [[x_{\alpha}(\xi), x_{\beta+\gamma}(\eta)], x_{\beta}(\zeta) ] & = 1, 
 & \xi,\eta,\zeta\in B_1,\ \text{ $(\alpha, \beta, \gamma)$ is an $\rA_3$-triple }.
\end{align}
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

The natural embedding of generators $\mathcal{X}_n \subseteq \mathcal{X}_{n+1}$ induces a map $\theta_n\colon G_n \to G_{n+1}$.
Our goal is to show the following result.
\begin{prop}\label{prop:tul3.3}
 For $n\geq 1$ the map $\theta_n$ is an isomorphism. Consequently, $\St(\Phi, B)$ can be presented by generators $\mathcal{X}_1$
  and families of relations $(a_1)$, $(b_1)$, $(c_1)$, $(d_1)$.
\end{prop}
%TODO: Deduce 2nd assertion from 1st
Notice that the second assertion follows from the first one.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.
%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)

@@

The following lemma is similar to~\cite[Proposition 1.1]{Re75} and~\cite[Proposition~3.2.2]{RS76}.
\begin{lemma}
 Suppose $m \geq 1$.
 Let $\alpha, \beta, \alpha', \beta'$ be such that $\alpha + \beta = \alpha' + \beta'$.
 Assume, moreover, that $\xi \in B_k$, $\xi' \in B_{k'}$, $\eta \in B_l$, $\eta' \in B_{l'}$ are such that 
  $N_{\alpha, \beta} \xi \eta = N_{\alpha', \beta'}\xi' \eta'$ for some $k,k',l,l'\leq m$ satisfying $k+l=k'+l' = m+1$.
 Then in the group $G_m$ the following relations hold:
 \begin{align}
  \label{item:i} [x_\alpha(\xi), x_\beta(\eta)] &= [x_{\alpha'}(\xi'), x_{\beta'}(\eta')] & \\
  \label{item:ii} [[x_\gamma(\zeta), [x_\alpha(\xi), x_\beta(\eta)]] & = 1 & \text{in the case } \gamma\in\{\alpha, \beta, \alpha + \beta\},\\
  \nonumber                                     &                          & \text{$\zeta \in B_{k''}$, $k''\leq m$.}
 \end{align}
\end{lemma}
\begin{proof}
 Notice that $k+l = m+1$, $k, l\leq m$ imply $k,l>0$, hence $B_i= t^i \cdot A$ for $i=k,k',l,'l'$.
 Thus, the argument of~\cite[Proposition 1.1]{Re75} can be repeated verbatim.
\end{proof}

\begin{lemma} For every $\rA_3$-triple $(\alpha, \beta, \gamma)$ in $\Phi$ the following relation holds in $G_1$:
\begin{align} 
 [[x_{\beta}(b t), x_{\alpha}(a t)],  x_{\gamma}(c)] & = [x_{\alpha}(at), x_{\beta+\gamma}(N_{\beta, \gamma}b c t)].
\end{align}
 \end{lemma}
\begin{proof}
 \begin{align*}
   [[x_{\beta}(b t), x_{\alpha}(a t)],  x_{\gamma}(c)] = [{}^{x_{\beta}(b t)}\!x_{\alpha}(at), [x_{\gamma}(c),x_{\beta}(b t)]] &  \text{ by~\eqref{eq:H1iii} and~$(b_1)$} \\ 
                                                       = {}^{x_{\beta}(b t)}[ x_{\alpha}(at), x_{\beta +\gamma}(N_{\beta,\gamma}bct) ] & \text{ by $(b_1)$ and $(c_1)$}\\ 
                                                       %TODO:  Missing structure constants...
                                                       = [x_{\alpha}(at), x_{\beta +\gamma}(N_{\beta,\gamma}bct)] & \text{ by $(d_1)$. \qedhere}
 \end{align*}

\end{proof}

To construct a map splitting to $\theta_m$ we construct the missing generators of $G_{m+1}$ in $G_m$ 
 and then show that they satisfy relations $a_{m+1}$, $b_{m+1}$ and $c_{m+1}$.

For every $\xi \in B_{m+1}$ and $\alpha\in \Phi$ there exist $\xi' \in B_m$ and $\alpha'\in \Phi$ such that $\xi = t\xi'$ and $\alpha-\alpha'\in\Phi$.
Thus, we can make the following definition:
\begin{equation} \label{eq:defmpn} x_\alpha(\xi) := [x_{\alpha-\alpha'}(N_{\alpha-\alpha',\alpha'} \xi'), x_{\alpha'}(t)],\end{equation} 
and from~\eqref{item:i} its correctness (i.\,e. the independence of the choice of $\alpha'$) follows.

We now turn to the verification of relations $(a_{m+1})$, $(b_{m+1})$, $(c_{m+1})$.
Notice first that \eqref{eq:H1ii} and~\eqref{item:ii} immediately imply $(a_{m+1})$ and hence $(b_{m+1})$ in the special case $\alpha=\alpha'$.

To verify relations $(c_{m+1})$ it suffices to show that
\begin{equation}
\label{eq:verify-bmp1} [x_\alpha(\xi), x_{\alpha'}(at^{m+1})] = [x_\alpha(t\xi), x_{\alpha'}(at^m)],\ a \in A,\ \xi \in B_k,\ k\leq 0.
\end{equation}
We can find root subsystem $\Psi \subseteq \Phi$ of type $\mathsf{A}_3$ containing the roots $\alpha, \alpha'$.
Choose basis $\{\alpha,\beta,\gamma\}$ of $\Psi$ so that $\alpha'=\beta$ and the Dynkin diagram of $\Psi$ looks as follows: HERE WAS A DIAGRAM

Let us verify~\eqref{eq:verify-bmp1}:
\begin{align*}
   [x_\alpha(\xi), x_\beta(at^{m+1})] = [x_\alpha(\xi), [x_{\beta + \gamma}(t), x_{-\gamma}(a't^m)]]&  \text{ by~\eqref{eq:defmpn} for a suitable $a'\in A$} \\ 
 = [x_{\alpha+\beta+\gamma}(\epsilon t\xi), {}^{x_{\beta+\gamma}(t)}x_{-\gamma}(a't^m)]             &  \text{ by~\eqref{eq:H1iii}, for $\epsilon=N_{\alpha, \beta+\gamma}$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t\xi), x_{-\gamma}(a't^m)]             &  \text{ by $(b_1)$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t^2\xi), x_{-\gamma}(a't^{m-1})]       &  \text{ by~\eqref{item:i} if $k=0$ or~\eqref{eq:cm} if $k <0$} \\
 %TODO: There's a mistake at this moment --- k = 0, m = 1 and the needed relation is not in G_1 -- :(
 = [[x_\alpha(t\xi), x_{\beta+\gamma}(t)], {}^{x_{\beta+\gamma}(t)} x_{-\gamma}(a't^{m-1})]         &  \text{ by~$(b_2)$, $(c_2)$ or by~\eqref{item:ii},\eqref{eq:defmpn} if $m=1$} \\
 = [x_\alpha(t\xi), x_\beta(at^m)]                                                                  &  \text{ by~\eqref{eq:H1iii}.}
\end{align*}

>>>>

We need one additional technical definition. We call an ordered triple of roots $(\alpha, \beta, \gamma) \in \Phi^3$ an {\it $\rA_3$-triple}
 if these vectors form a basis of a root subsystem $\Psi \subseteq \Phi$ of type $\rA_3$, and moreover if the Dynkin diagram of $\Psi$ with respect to this basis has the form
 $\xymatrix{\circ_\alpha \ar@{-}[r] & \circ_\beta \ar@{-}[r] & \circ_\gamma}.$ %TODO: Draw better diagram :)

@@

For $m\geq 2$ define $G_m$ as the group presented by generators $\mathcal{X}_m$ and the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{\alpha}(\xi) x_{\alpha}(\eta) & = x_{\alpha}(\xi+\eta),&  \xi,\eta\in B_d,\ d\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_d,\ \eta \in B_e,\ d,e\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_d,\ \eta \in B_e,\ d, e, d+e \leq m.
 \end{align}
For $m=1$ define $G_1$ as the group presented by generators $\mathcal{X}_1$, 
three above families of relations $(a_1), (b_1), (c_1)$ and the following additional family of relations:
\begin{align}  \label{eq:d1} \tag{$d_1$} [[x_{\alpha+\beta+\gamma}(\xi), x_{-\gamma}(\eta)], x_{\beta+\gamma}(\zeta) ] & = 1, 
 & \xi,\eta,\zeta\in B_1,\ \text{ $(\alpha, \beta, \gamma)$ is an $\rA_3$-triple. }
\end{align}
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

The obvious embedding of generators $\mathcal{X}_m \subseteq \mathcal{X}_{m+1}$ induces a map $f_m\colon G_m \to G_{m+1}$.
Our goal is to show the following result.
\begin{prop}\label{prop:tul3.3}
 For $m\geq 1$ the map $f_m$ is an isomorphism. Consequently, $\St(\Phi, B)$ can be presented by generators $\mathcal{X}_1$
  and relations $(a_1)$, $(b_1)$, $(c_1)$, $(d_1)$.
\end{prop}
%TODO: Deduce 2nd assertion from 1st
Notice that the second assertion follows from the first one.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.
%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)

@@

The following lemma is similar to~\cite[Proposition 1.1]{Re75} and~\cite[Proposition~3.2.2]{RS76}.
\begin{lemma}
 Suppose $m \geq 1$.
 Let $\alpha, \beta, \alpha', \beta' \in \Phi$ be such that $\alpha + \beta = \alpha' + \beta'$.
 Assume, moreover, that $\xi \in B_k$, $\xi' \in B_{k'}$, $\eta \in B_l$, $\eta' \in B_{l'}$ are such that 
  $N_{\alpha, \beta} \xi \eta = N_{\alpha', \beta'}\xi' \eta'$ for some $k,k',l,l'\leq m$ satisfying $k+l=k'+l' = m+1$.
 Then the following relations holds in $G_m$:
 \begin{equation}
  \label{item:i} [x_\alpha(\xi), x_\beta(\eta)] = [x_{\alpha'}(\xi'), x_{\beta'}(\eta')].
 \end{equation}
 Moreover for every $\zeta \in B_{k''}$, $k''\leq m$ and $\gamma\in\{\alpha, \beta, \alpha + \beta\}$
  the following relation holds in $G_m$:
 \begin{equation}
 \label{item:ii} [[x_\gamma(\zeta), [x_\alpha(\xi), x_\beta(\eta)]] = 1.
 \end{equation}
\end{lemma}
\begin{proof}
 Notice that $k+l = m+1$, $k, l\leq m$ imply $k,l>0$, hence $B_i= t^i \cdot A$ for $i=k,k',l,'l'$.
 Thus, the argument of~\cite[Proposition 1.1]{Re75} can be repeated verbatim in our situation.
 %TODO: Write down this argument, at least a part of it?
\end{proof}

For every $\xi \in B_{m+1}$ and $\alpha\in \Phi$ there exist $\xi' \in B_m$ and $\alpha'\in \Phi$ such that $\xi = t\xi'$ and $\alpha-\alpha'\in\Phi$,
so we can define the following element of $G_m$:
\begin{equation} \label{eq:defmpn} S_\alpha(\xi) := [x_{\alpha-\alpha'}(N_{\alpha-\alpha',\alpha'} \xi'), x_{\alpha'}(t)].\end{equation} 
From~\eqref{item:i} it follows that $S_\alpha(\xi)$ does not depend of the choice of $\alpha'$.

%To show that $f_m$ is an isomorphism we construct a map $g_m \colon G_{m+1} \to G_m$ in the opposite direction.
%We define it on generators $\mathcal{X}_{m+1}$ and then verify that it is well defined.
%Clearly, $g_m$ should be defined as the identity map on the subset of generators $\mathcal{X}_m \subseteq \mathcal{X}_{m+1}$.
%Now if $x_{\alpha}(\xi)$ is an element of $\mathcal{X}_{m+1}\setminus \mathcal{X}_m$
%there exist $\xi' \in B_m$ and $\alpha'\in \Phi$ such that $\xi = t\xi'$ and $\alpha-\alpha'\in\Phi$, 
% therefore we can set 
%\[g_m(x_\alpha(\xi)) = [x_{\alpha-\alpha'}(\xi'), x_{\alpha'}(N_{\alpha-\alpha',\alpha'} t)].\]
%It follows from~\eqref{item:i} that this definition does not depend on the choice of $\alpha'$.

Let us show that elements $S_\alpha(\xi)$ satisfy the relations $(a_{m+1})$, $(b_{m+1})$ and $(c_{m+1})$.
From \eqref{eq:H1ii} and~\eqref{item:ii} it follows that $S_\alpha(\xi)$  satisfy $(a_{m+1})$ and hence $(b_{m+1})$ in the special case $\alpha=\alpha'$.
\begin{comment}
\begin{lemma} For every $\rA_3$-triple $(\alpha, \beta, \gamma)$ in $\Phi$ and $a,b,c \in A$ the following relation holds in $G_1$:
\begin{align} 
 [x_{\alpha}(a), [x_{\beta}(b t), x_{\gamma}(c t)]] & = [x_{\beta+\alpha}(N_{\beta, \alpha}a b t), x_{\gamma}(ct)].
\end{align}
 \end{lemma}
\begin{proof}
 \begin{align*}
   [x_{\alpha}(a), [x_{\beta}(b t), x_{\gamma}(c t)]] = [[x_{\alpha}(a), x_{\beta}(b t)], {}^{x_{\beta}(b t)}\!x_{\gamma}(ct)] &  \text{ by~\eqref{eq:H1iii} and~$(b_1)$} \\ 
                                                       = {}^{x_{\beta}(b t)}[x_{\alpha +\beta}(N_{\alpha,\beta}abt), x_{\gamma}(ct) ] & \text{ by $(b_1)$ and $(c_1)$}\\ 
                                                       = [x_{\alpha +\beta}(N_{\alpha,\beta}abt), x_{\gamma}(ct)] & \text{ by $(d_1)$. \qedhere}
 \end{align*}  
\end{proof}
\end{comment}

\begin{lemma} Suppose that $m\geq 1$. For every $\alpha, \alpha' \in \Phi$ such that $\alpha+\alpha' \in \Phi$ and
 $a\in A$, $\xi \in B_k$, $k \leq 0$ the following relation holds in $G_m$: 
\begin{equation}
\label{eq:verify-bmp1} [x_\alpha(\xi), S_{\alpha'}(at^{m+1})] = [x_\alpha(t\xi), x_{\alpha'}(at^m)].
\end{equation}
\end{lemma}
\begin{proof}
By \cite[Lemma~3.1.2]{RS76} we can assume without loss of generality that $\alpha'=\beta$ for some $\rA_3$-triple $(\alpha, \beta, \gamma)$.
\begin{align*}
   [x_\alpha(\xi), S_\beta(at^{m+1})] = [x_\alpha(\xi), [x_{\beta + \gamma}(t), x_{-\gamma}(a't^m)]]&  \text{ by~\eqref{item:i},\eqref{eq:defmpn} for $a' = N_{\beta+\gamma, -\gamma} a$} \\ 
 = [x_{\alpha+\beta+\gamma}(\epsilon t\xi), {}^{x_{\beta+\gamma}(t)}x_{-\gamma}(a't^m)]             &  \text{ by~\eqref{eq:H1iii}, for $\epsilon=N_{\alpha, \beta+\gamma}$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t\xi), x_{-\gamma}(a't^m)]             &  \text{ by $(b_1)$}
\end{align*} 
There are two cases.
\begin{enumerate}
 \item \label{case:cm-1} Case $m=1$, $k = 0$.
 \begin{align*}
   RHS  = [x_{\alpha +\beta + \gamma}(\epsilon t\xi), x_{-\gamma}(a't^m)] & \text{ by $(d_1)$.}\\
        = [x_\alpha(t\xi), x_\beta(at^m)]                                 & \text{ by~\eqref{item:i}. }   
 \end{align*}  
 This uses $N_{\alpha, \beta+\gamma} N_{\alpha+\beta+\gamma, -\gamma} N_{\beta+\gamma, -\gamma} = N_{\alpha, \beta}$. Why is this true?

 \item \label{case:cm-2} Case $m\geq 2$ or $m=1$, $k < 0$.
 \begin{align*}
 RHS = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t^2\xi), x_{-\gamma}(a't^{m-1})]       &  \text{ by~\eqref{item:i} if $k=0$ or~\eqref{eq:cm} if $k <0$} \\
 %TODO: There's a mistake at this moment --- k = 0, m = 1 and the needed relation is not in G_1 -- :(
 = [[x_\alpha(t\xi), x_{\beta+\gamma}(t)], {}^{x_{\beta+\gamma}(t)} x_{-\gamma}(a't^{m-1})]         &  \text{ by~$(b_2)$, $(c_2)$ or by~\eqref{item:ii},\eqref{eq:defmpn} if $m=1$} \\
 = [x_\alpha(t\xi), x_\beta(at^m)]                                                                  &  \text{ by~\eqref{eq:H1iii}.}
\end{align*}

\end{enumerate}
\end{proof}
